Quantum Algorithms for Portflolio Optimisation in Finance

### 2024-05-29 project/paper-v2.tex
update project paper

Then, the first information we extract from the data is a list $P$ of current prices $P_i$  of the considered assets.

\begin{equation}
  \label{eq:1}
  P_i = p_i(T)
\end{equation}

In addition, the historical returns of the assets between time $t-1$ and $t$ are calculated as:

\begin{equation}
  \label{eq:2}
  r_i(t) = \frac{p_i(t) - p_i(t-1)}{p_i(t-1)}
\end{equation}

But for inference purposes, we must consider the expected return of the assets. So, given the historical returns, we can calculate the expected return of the assets as:

\begin{equation}
  \label{eq:3}
  \mu_i = \frac{1}{T} \sum_{t=1}^{T} r_i(t)
\end{equation}

The variance of the returns of each asset and the covariance between returns of different assets over the historical period are given by:

\begin{equation}
  \label{eq:4}
  \sigma_i^2 = \frac{1}{T-1} \sum_{t=1}^{T} (r_i(t) - \mu_i)^2
\end{equation}

\begin{equation}
  \label{eq:5}
  \sigma_{ij} = \frac{1}{T-1} \sum_{t=1}^{T} (r_i(t) - \mu_i)(r_j(t) - \mu_j)
\end{equation}

@@

>>>>

Then, the first information we extract from the data is a list $P$ of current prices $P_i$  of the considered assets.
\begin{equation}
  \label{eq:1}
  P_i = p_i(T)
\end{equation}
In addition, the historical returns of the assets between time $t-1$ and $t$ are calculated as:
\begin{equation}
  \label{eq:2}
  r_i(t) = \frac{p_i(t) - p_i(t-1)}{p_i(t-1)}
\end{equation}
But for inference purposes, we must consider the expected return of the assets. So, given the historical returns, we can calculate the expected return of the assets as:
\begin{equation}
  \label{eq:3}
  \mu_i = \frac{1}{T} \sum_{t=1}^{T} r_i(t)
\end{equation}
The variance of the returns of each asset and the covariance between returns of different assets over the historical period are given by:
\begin{equation}
  \label{eq:4}
  \sigma_i^2 = \frac{1}{T-1} \sum_{t=1}^{T} (r_i(t) - \mu_i)^2
\end{equation}
\begin{equation}
  \label{eq:5}
  \sigma_{ij} = \frac{1}{T-1} \sum_{t=1}^{T} (r_i(t) - \mu_i)(r_j(t) - \mu_j)
\end{equation}
A set of investments $x_i$ (given as a fraction of the total budget or number of asset units) allocated for each $i$th considered asset, 
is then defined. Therefore, an optimal method for asset allocation aims to maximise the portfolio return 
$\mu^\text{T} x$ while minimising risk, defined as the portfolio variance $x^\text{T}\Sigma x$, where $\mu$ is the vector of expected returns for each asset $i$ computed by Eq. \ref{eq:3}, $\Sigma$ is the covariance matrix calculate byt Eq. \ref{eq:4} and Eq. \ref{eq:5}, and $x$ is the vector 
of asset allocations measured as fractions of the total budget. Therefore, the problem of finding the optimal portfolio is reduced to find the $x$ vector that maximises the following objective function: 
\begin{equation}
  \label{eq:6}
  \mathcal{L}(x) = \mu^\text{T} x - q x^\text{T}\Sigma x
\end{equation}
where $q$ is a parameter that determines the trade-off between risk and return.  In a realistic scenario, budget constraints and non-negative 
$x_i$ values (only buying) must be considered. Therefore, in the general case, the problem is formulated as follows:
\begin{equation}
\begin{aligned}
  \label{eq:7}
  \max_{x} \mathcal{L}(x): & \max_{x}\left(\mu^\text{T} x - q x^\text{T}\Sigma x \right) \\ 
  \text{s.t.} & \quad \sum_{i=1}^{N} x_i = 1 
\end{aligned}
\end{equation}
In section \ref{sec:literature1} we will review how this problem can be mapped into a QUBO problem and solved using quantum algorithms.

@@

An option is finan
